JavaScript as an Embedded DSL

### 2012-03-29 papers/ecoop2012/article.tex
move the discussion of our contributions to LMS to the end of the intro

We propose to embed JavaScript as a DSL in a host language.\footnote{Surely, the embedded language is not exactly JavaScript: as an embedded language, it naturally is a subset of Scala. However it is quite close to JavaScript: often one can take snippets of JavaScript code and use them in the DSL with minor syntactic tweaking (as demonstrated by the Snowflake example).} Through LMS, we tackle the challenges outlined above with minimal effort, as most of the work is off-loaded to the host language. In particular, we make the following contributions:
\begin{itemize}
\item Our DSL is statically typed through the host language, yet supports gradual typing notably for incorporating external JavaScript libraries and APIs (section~\ref{sec:typing}).
\item In addition to generating JavaScript code, our DSL can be executed directly in the host language, allowing code to be shared between client and server (section~\ref{sec:trivial-embedding}).
\item We use advanced object-oriented techniques to achieve modularity in our DSL: each language primitive and API is defined in a separate module (section~\ref{sec:modularity}).
\item Our DSL supports typed object literals and class-based objects. The translations to JavaScript are lightweight and intuitive: the object literals translate to JSON-like object literals and the class-based objects to JavaScript constructor-based objects (section~\ref{sec:reification}).
\item With minimal effort, we exploit the selective CPS transform already existing in the host language to provide a compelling abstraction over asynchronous callback-driven programming in our DSL (section~\ref{sec:cps}). This case-study demonstrates the fruitfulness of re-using host language features to enhance our embedded DSL. A novel insight is that CPS transforming a program generator allows it to generate code that is in CPS.
\item The present work extends both the range of virtualizable language features and the LMS framework, which is beneficial to future DSL efforts taking the JavaScript work as a case study. Previous LMS embeddings had to define each operation explicitly (like in section~\ref{sec:defineDslComponent}). This paper contributes lifting of whole traits or classes (through the \code{repProxy} mechanism described and used in sections~\ref{sec:typedApis}\&~\ref{sec:reified-classes}), untyped or optionally typed operations (sections~\ref{sec:casting},~\ref{sec:dynamic}\&~\ref{sec:dynamic2static}), and typed object literals (section~\ref{sec:object-literals}) including the necessary Scala-Virtualized support.
\end{itemize}

@@

>>>>

We propose to embed JavaScript as a DSL in a host language.\footnote{Surely, the embedded language is not exactly JavaScript: as an embedded language, it naturally is a subset of Scala. However it is quite close to JavaScript: often one can take snippets of JavaScript code and use them in the DSL with minor syntactic tweaking (as demonstrated by the Snowflake example).} Through LMS, we tackle the challenges outlined above with minimal effort, as most of the work is off-loaded to the host language. In particular, we make the following contributions:
\begin{itemize}
\item Our DSL is statically typed through the host language, yet supports gradual typing notably for incorporating external JavaScript libraries and APIs (section~\ref{sec:typing}).
\item In addition to generating JavaScript code, our DSL can be executed directly in the host language, allowing code to be shared between client and server (section~\ref{sec:trivial-embedding}).
\item We use advanced object-oriented techniques to achieve modularity in our DSL: each language primitive and API is defined in a separate module (section~\ref{sec:modularity}).
\item Our DSL supports typed object literals and class-based objects. The translations to JavaScript are lightweight and intuitive: the object literals translate to JSON-like object literals and the class-based objects to JavaScript constructor-based objects (section~\ref{sec:reification}).
\item With minimal effort, we exploit the selective CPS transform already existing in the host language to provide a compelling abstraction over asynchronous callback-driven programming in our DSL (section~\ref{sec:cps}). This case-study demonstrates the fruitfulness of re-using host language features to enhance our embedded DSL. A novel insight is that CPS transforming a program generator allows it to generate code that is in CPS.
\end{itemize}

@@

The present work extends both the range of virtualizable language features and the LMS framework, which is beneficial to future DSL efforts taking the JavaScript work as a case study. Previous LMS embeddings had to define each operation explicitly (like in section~\ref{sec:defineDslComponent}). This paper contributes lifting of whole traits or classes (through the \code{repProxy} mechanism described and used in sections~\ref{sec:typedApis}\&~\ref{sec:reified-classes}), untyped or optionally typed operations (sections~\ref{sec:casting},~\ref{sec:dynamic}\&~\ref{sec:dynamic2static}), and typed object literals (section~\ref{sec:object-literals}) including the necessary Scala-Virtualized support.
